---
title: "p8105_hw5_jm5916"
author: "Juhi Mattekatt"
date: "2025-11-13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

set.seed(1)
```
# Problem 1
### Create function
```{r birth_function}
birthday = function(x) {
  birthdays = sample(1:365, x, replace = TRUE) #select bdays (equal distribution), allow for repeats
  return(any(duplicated(birthdays))) #true if there are any repeated bdays
}

#test function
birthday(5)
```

### Simulate groups and plot
```{r birth_sim}
#Simulation
probabilities <- tibble(
  group_size = 2:50,
  probability = map_dbl(2:50, ~mean(replicate(10000, birthday(.))))
)

#Plot
probabilities %>% 
ggplot(aes(x=group_size, y=probability)) + 
  geom_bar(stat = "identity", fill = "thistle3") +
  labs(title = "Birthday Groups",
       x = "Number of people in group",
       y = "Probability of shared birthday") +
  theme_minimal()

```

The probability of two people sharing a birthday increases as the number of people increases. It goes over 50% around 23 people, and the probability is over 90% at 40 people.

# Problem 2
### Create function
```{r setup_p2}
sim_mean_sd = function(n = 30, mu = 0, sigma = 5) { #fix conditions
  
  sample = rnorm(n, mean = mu, sd = sigma) #normal distribution
  
   results = t.test(sample, mu = 0) %>% 
     broom::tidy() %>% 
     select(estimate, p.value) #inlcude only estimate and p-value
   
   return(results)
}
```

### Simulation
```{r simulation}
#for just mu = 0
sim_results_0 = 
  map(1:5000, ~ sim_mean_sd(mu = 0)) %>% #run 5000 times
  bind_rows()

#simulation for all mu's
sim_results_list = list()
mu_values = c(0, 1, 2, 3, 4, 5, 6)

for(mu in mu_values) { #using a for loop this time to practice 
  sim_results <- map(1:5000, ~ sim_mean_sd(mu = mu)) %>%
    bind_rows() %>% 
    mutate(mu=mu) #replace mu for every new loop
  
  sim_results_list[[as.character(mu)]] = sim_results #list contains all samples for each mu
}
```

### Power Plot
```{r null_plot}
#combine results into one df for plots
sim_results_all = bind_rows(sim_results_list, .id = "mu")

#create power plot
power = sim_results_all %>%
  group_by(mu) %>%
  summarise(total_sim = n(),
            reject = sum(p.value < 0.05),
            power = reject/total_sim) %>% #power equation
  ggplot(aes(x = as.numeric(mu), y = power)) +
  geom_line() +
  geom_point() +
  labs(title = "Power of Test",
       x = "True value of μ",
       y = "Proportion of p-value < 0.05") +
  theme_minimal()

power
```

Proportion of p-value < 0.05, or power, increases as the true value of μ increases, this suggests that power increases with effect size assuming that sample size and sd remain constant. 

### Estimate Plot
```{r est_plot}
estimate <- sim_results_all %>%
  group_by(mu) %>%
  summarise(
    av_est = mean(estimate), #average estimate of 𝜇
    av_null_est = mean(estimate[p.value < 0.05]), #average estimate of 𝜇̂ only in samples for which the null was rejected
    reject_count = sum(p.value < 0.05), #check to make sure I have correct numbers
    .groups = "drop") %>%
  ggplot(aes(x = as.numeric(mu))) +
  geom_line(aes(y = av_est, color = "All")) + 
  geom_point(aes(y = av_est, color = "All"), shape = "circle") +
  geom_line(aes(y = av_null_est, color = "Only null rejected")) + #overlaying plots
  geom_point(aes(y = av_null_est, color = "Only null rejected"), shape = "square") +
  labs(
    title = "Estimate to True μ",
    x = "True value of μ",
    y = "Average estimate of μ",
    color = "Tests included"
  ) +
  scale_color_manual(values = c("All" = "sienna2", "Only null rejected" = "chartreuse4")) + 
  theme_minimal()

estimate
```

The green line representing sample average for tests in which the null was rejected deviates away from the line representing the sample average of all the tests at smaller true values of mu and comes closer to the line at bigger values of mu. This suggests that the sample average of estimated 𝜇 across tests for which the null is rejected is approx. equal to the true value of 𝜇 only when the effect size is large. 

This is likely because removing samples from our simulation as we do when only including those in which the null is rejected introduces bias to our estimate. Thus, the more samples we remove, the larger our bias and the more our estimate deviates away from the truth. This is especially apparent in conditions where the effect size is small because these estimates have smaller power and a greater number of their samples will be removed. Larger effect sizes don't have this problem because their tests tend to have more power and fewer of their samples are removed, leading to less bias (there is still some bias however as we are still only including those tests where the estimate was large enough to have enough power to reject the null hypothesis).

# Problem 3
### Import data from web
```{r import_web_data, message = FALSE}
library(readr)
homicides = 
  read_csv("http://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv")
#viewed data in console
```
This dataset from the Washington Post lists homicide data from 50 major US cities. It has `r nrow(homicides)` observations and `r ncol(homicides)` columns, and includes information like victim name, date of report, race, age, and city. 

### Create new variable city_state
```{r city_state}
city_state_hom = homicides %>% 
  mutate(
    city_state = paste(city, state, sep = ", ")
  ) %>% 
  group_by(city_state) %>% 
  summarise(total_hom = n(),
            unsolved = sum(disposition %in% c("Closed without arrest", "Open/No arrest")))

#view in console
```
There are `r nrow(city_state_hom)` cities listed here. There is an observation where Tulsa was listed as being in AL, which might explain this "extra" city. I will not fix this mistake but I will drop this observation for future steps. 

### Estimated proportion of unsolved cases in Baltimore, MD
```{r baltimore_proptest, warnings = FALSE}
baltimore_hom = city_state_hom %>% 
  filter(city_state == "Baltimore, MD")

baltimore_hom_test = 
  prop.test(x = pull(baltimore_hom, unsolved),
          n = pull(baltimore_hom, total_hom)) %>% 
  broom::tidy() %>% 
  select(estimate, conf.low, conf.high)

baltimore_hom_test
  
# estimated proportion of unsolved homicides = 0.65 (95% CI: 0.63, 0.66)
```

### Estimated proportion of unsolved cases for all cities
```{r all_cities_proptest}
city_prop <- city_state_hom %>% 
  filter(city_state != "Tulsa, AL") %>% #remove miscode
  mutate(prop_test_results = map2(unsolved, total_hom, ~ prop.test(.x, .y) %>% #use map2 to repeat test.prop
                                    broom::tidy())) %>% #tidy 
  unnest(prop_test_results) %>% #make results accessible 
  select(city_state, estimate, conf.low, conf.high) #only want estimated proportions and CIs for each city

head(city_prop)
```

#### Plot
```{r murder_plot, fig.height = 12, fig.width = 9}
city_prop %>% 
  arrange(estimate) %>% 
  mutate(city_state = fct_inorder(city_state)) %>% 
  ggplot() + 
geom_errorbar(aes(y=city_state, xmin=conf.high, xmax=conf.low), width = 0.3, color = "steelblue4") + 
geom_point(aes(y=city_state, x=estimate), size = 1.5, color = "steelblue3") +
  labs(
    title = "Estimated proportion of unsolved homicide cases in 50 major US cities",
    x = "Proportion Unsolved",
    y = "City") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, hjust = 1, lineheight = 1.5))
```

Chicago, IL, New Orleans, LA, and Baltimore, MD have the highest estimated proportion of unsolved homicide cases among the cities included in this dataset. No city has more than 75% of homicides unsolved. Richmond, VA has the smallest estimated proportion of unsolved cases at just over 25%. Confidence intervals vary with sample size.
