---
title: "p8105_hw5_jm5916"
author: "Juhi Mattekatt"
date: "2025-11-13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(ggplot2)

set.seed(1)
```
# Problem 1

Suppose you put 𝑛 people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.
```{r birth_function}
birthday = function(x) {
  birthdays = sample(1:365, x, replace = TRUE) #select bdays (equal distribution), allow for repeats
  return(any(duplicated(birthdays))) #true if there are any repeated bdays
}
```

Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. Make a plot showing the probability as a function of group size, and comment on your results.
```{r birth_sim}
n = 2:50
probabilities = tibble(
  group_size = 2:50,
  probability = sapply(n, function(n) {
  mean(replicate(10000, birthday(n)))
}) 
)

probabilities %>% 
ggplot(aes(x=group_size, y=probability)) + 
  geom_bar(stat = "identity")

```

The probability of two people sharing a birthday increases as the number of people increases. It goes over 50% around 23 people, and the probability is over 90% at 40 people.

# Problem 2
### Create Function
```{r setup_p2}
sim_mean_sd = function(n = 30, mu = 0, sigma = 5) { #fix conditions
  
  sample = rnorm(n, mean = mu, sd = sigma) #normal distribution
  
   results = t.test(sample, mu = 0) %>% 
     broom::tidy() %>% 
     select(estimate, p.value) #inlcude only estimate and p-value
   
   return(results)
}
```

### Simulation
```{r simulation}
#for just mu = 0
sim_results_0 = 
  map(1:5000, ~ sim_mean_sd(mu = 0)) %>% #run 5000 times
  bind_rows()

#simulation for all mu's
sim_results_list = list()
mu_values = c(0, 1, 2, 3, 4, 5, 6)

for(mu in mu_values) { #using a for loop this time
  sim_results <- map(1:5000, ~ sim_mean_sd(mu = mu)) %>%
    bind_rows() %>% 
    mutate(mu=mu) #replace mu for every new loop
  
  sim_results_list[[as.character(mu)]] = sim_results #list contains all samples for each mu
}
```

### Power Plot
```{r null_plot}
#combine results into one df for plots
sim_results_all = bind_rows(sim_results_list, .id = "mu")

#create power plot
power = sim_results_all %>%
  group_by(mu) %>%
  summarise(total_sim = n(),
            reject = sum(p.value < 0.05),
            power = reject/total_sim) %>% #power equation
  ggplot(aes(x = as.numeric(mu), y = power)) +
  geom_line() +
  geom_point() +
  labs(title = "Power of Test",
       x = "True value of μ",
       y = "Proportion of p-value < 0.05") +
  theme_minimal()

power
```

### Estimate Plot
```{r est_plot}
estimate <- sim_results_all %>%
  group_by(mu) %>%
  summarise(
    av_est = mean(estimate), #average estimate of 𝜇
    av_null_est = mean(estimate[p.value < 0.05]), #average estimate of 𝜇̂ only in samples for which the null was rejected
    reject_count = sum(p.value < 0.05), #check to make sure I have correct numbers
    .groups = "drop") %>%
  ggplot(aes(x = as.numeric(mu))) +
  geom_line(aes(y = av_est, color = "All")) + 
  geom_point(aes(y = av_est, color = "All"), shape = "circle") +
  geom_line(aes(y = av_null_est, color = "Only null rejected")) + #overlaying plots
  geom_point(aes(y = av_null_est, color = "Only null rejected"), shape = "square") +
  labs(
    title = "Estimate to True μ",
    x = "True value of μ",
    y = "Average estimate of μ",
    color = "Tests included"
  ) +
  scale_color_manual(values = c("All" = "sienna2", "Only null rejected" = "chartreuse4")) + 
  theme_minimal()

estimate
```

The green line representing sample average for tests in which the null was rejected deviates away from the line representing the sample average of all the tests at smaller true values of mu and comes closer to the line at bigger values of mu. This suggests that the sample average of estimated 𝜇 across tests for which the null is rejected ia approx. equal to the true value of 𝜇 only when the effect size is large. This is likely because removing samples from our simulation as we do when only including those in which the null is rejected introduces bias to our estimate. Thus, the more samples we remove, the larger our bias and the more our estimate deviates away from the truth. This is especially apparent in conditions where the effect size is small because these estimates have smaller power and a greater number of their samples will be removed, larger effect sizes don't have this problem because they tend to have more power and fewer of their samples were removed, leading to less bias (there is still some bias however as we are still only including those tests where the estimate was large enough to have enough power to reject the null hypothesis).

# Problem 3

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

